---
title: "Neo4j and R: A network of possibilities"
author: "MAJ Nicholas Uhorchak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_notebook
---
## Introduction
Self described as the "World's leading graph database, with native graph storage and processing," [neo4j](https://neo4j.com/) is a graph database, with its own native query language, Cypher. Neo4j is extremely flexible, as both nodes and relationships contain attributes in the form of JSON documents.   

Neo4j provides great [tutorials](https://neo4j.com/graphacademy/online-training/), as well as [worst practices](https://neo4j.com/blog/dark-side-neo4j-worst-practices/) guide from their website.  Although these tutorials are great references, users shouls familiarize themselves with both database and graph network basics prior to delveing deep into this demo.

## Creating a database in R

```{r}
require(neo4r)
require(magrittr)
```

- Please note, that Neo4j native connection protocol is [*bolt*](https://boltprotocol.org/) and the desktop client will automatically launch a web browser that way.  The *current* neo4j driver package [(neo4R)](https://github.com/neo4j-rstats/neo4r) for R does not currently support bolt.

For this project, a blank database was created using the neo4j desktop client (add graph function from the neo4j GUI).  Once created, the 'Start' button must be clicked to move the database to an active state.

- Although provided for this demo, it is good practice not to store database names and passwords in public documents.
```{r}
# Connect to Neo4J
 
 con <- neo4j_api$new(url = 'http://localhost:7474/', user = 'neo4j', password = 'markdown')
con$ping() 
launch_con_pane(con)
```


need to create a database object, and connect to that database to run this query.

```{r}
match(n:Branch),(m:Division)
where n.Name =~ 'J6S.*' and m.Name = 'J6S'
create(n)-[r:works_for]->(m)
return(n),(m)
```

```{r}
create(J6:Directorate {Name:'J6', Title:'Chief Information Office'})
```

```{r}
MATCH (n { Name: 'J636' })
SET n.Title = toString('Satellite, Spectrum and Terrestrial')
RETURN n.name, n.Title
```

